---
title: "Basic Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r}
library(EffortDataTools)
library(gapindex)
```

## Establish an Oracle connection

### Database Authentication

Functions querying Oracle tables directly rely on an active database connection established via `gapindex`. You can either pass an existing connection object via the channel parameter or allow functions to connect automatically.

```{r}
con <- gapindex::get_connected(check_access = FALSE)
```

## Getting Haul Data for Repicks

To export haul metadata directly to a CSV file for repicks, use `pull_repick_data()`:

```{r}
repick_df <- pull_repick_data(
  cruise      = "202601",
  vessel      = "176",
  channel     = con,
  output_path = "haul_data_202601_176.csv"
)
```

## Plot Haul Track over Station Grids

To visualize a haul's GPS track overlaid on survey grid cells shaded by depth stratum, call `plot_haul_track()`:

```{r}
haul_spatial <- plot_haul_track(
  cruise  = 202601,
  vessel  = 176,
  haul    = 194,
  region  = "ai",
  channel = con
)

# Access spatial features returned by the function
track_sf <- haul_spatial$track
grid_sf  <- haul_spatial$grid
```

## Merging Multiple XML Files

During survey operations, the Scientific Computing System (SCS) logs may end up split across multiple sequential XML files. The `merge_scs_files()` function combines these XML logs into a single file for upload to Poseidon, orders events chronologically, and allows users to crop data outside active tow windows.

```{r}
xml_files <- c(
  "data/SCS_Log_001.xml",
  "data/SCS_Log_002.xml",
  "data/SCS_Log_003.xml"
)

# Merge XML files into a single master document
merged_doc <- merge_scs_files(
  file_paths  = xml_files,
  output_path = "output/Merged_Haul.xml"
)
```

#### Time-Trimming Records

You can optionally supply start_time and end_time parameters in local time (e.g., Alaska Daylight Time). The function automatically converts local timestamps to UTC and crops timestamped entries occurring outside the active tow window:

```{r}
merged_trimmed <- merge_scs_files(
  file_paths  = xml_files,
  output_path = "output/Haul0073_Trimmed.xml",
  start_time  = "2026-06-22 12:34:00",
  end_time    = "2026-06-22 13:46:00",
  tz          = "America/Anchorage"
)
```

# Data Finalization

## Checking Depth Errors

The `check_haul_depths()` function scans unfinalized hauls for missing gear or bottom depths (0) or cases where gear depth meets or exceeds bottom depth:

```{r}
# Audit depth records for the 2026 Aleutian Islands survey
depth_errors <- check_haul_depths(
  cruise  = 202601,
  region  = "AI",
  channel = con
)

head(depth_errors)
```

## Evaluating Hauls for Abundance

The `check_haul_abundance()` function joins unfinalized catch and haul records to flag hauls where abundance should be false (e.g., non-standard gear or accessories, bad performance codes, short tow duration relative to catch weight, or duplicate stations):

```{r}
# Flag operational haul issues
haul_issues <- check_haul_abundance(
  cruise  = 202601,
  region  = "AI",
  channel = con
)

head(haul_issues)
```
